\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-03-16 doc/paper.Rmd
up to chumbley score

Comparisons of striae from bullets are usually based on comparisons of striae in land engraved areas, which are extracted in form of cross sections, called *profiles* \citep{aoas,ma2004}. From profiles bullet *signatures* \citep{chu2013,aoas} are extracted as residuals of a loess fit (or Gaussian filter).
 They are considered to be noise free and a good reflection of the key attributes of the raw marking, and the unique features of a bullet. \hh{Do you have a reference for the previous sentenced?} A  detailed discussion of the extraction technique for signatures is given in \citet{aoas}.
 <!--, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.-->

\hh{don't split the discussion on the size. between the next paragraph and }
<!--The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}.--> The scans for the Hamby study come from two sources namely, NIST scans \citep{nist} and scans produced at Iowa State University \hh{Iowa State is too general. it's the High-resolution microscopy lab and we are CSAFE}. Both scans were generated using confocal light microscopes at a 20x magnification but the microscopes collected data at different resolutions. The NIST database contains scans of the bullet sets Hamby 44 and 252 at a resolution of 1.5625 $\mu m$ per pixel. The Iowa State scans are available for only the Hamby 44 bullet set and at a resoultion of 0.645 $\mu m$ per pixel. The data used by \citet{chumbley, hadler} on the other hand, was produced by a profilometer at a resolution of about 0.73 $\mu m$ per pixel. The different resolutions indicate that the length of the digitized markings would be significantly different for each setting.

@@

Markings made by screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length. The Ruger pistol barrels used in the  Hamby study have a 9mm caliber. This means that bullets used in these pistols have land markings of only about a quarter of the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising the problem of potential loss in power of the test. 
\hh{The Chumbley score allows a separation of a toolmark into raw and normalized digitized versions. Originally, this mechanism is intended to separate between class characteristics and individual markings, however, in the setting of matching bullet striae, we could also use it to separate bullet curvature in profiles from signatures before matching signatures. }

>>>>

Comparisons of striae from bullets are usually based on comparisons of striae in land engraved areas, which are extracted in form of cross sections, called *profiles* \citep{aoas,ma2004}. From profiles bullet *signatures* \citep{chu2013,aoas} are extracted as residuals of a loess fit or Gaussian filter.
 Signatures are considered to be noise free and a good reflection of the key attributes of the raw marking, and the unique features of a bullet. \hh{Do you have a reference for the previous sentenced?} A  detailed discussion of the extraction technique for signatures is given in \citet{aoas}.
 <!--, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.-->

\hh{don't split the discussion on the size. between the next paragraph and }
<!--The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}.--> 
There are two sources of scans for sets from the Hamby study available to us: scans of Hamby 44 and Hamby 252 are available from the NIST database \citep{nist}. Hamby 44 has also been made available to us and has been scanned locally for CSAFE at the Roy J.\ Carver High Resolution Microscopy Facility using a Sensofar confocal light microscope. 
Scans in the NIST database are made with a NanoFocus at 20x magnification. The resolutions of the two instruments are different: the NIST scans are taken at a resolution of 1.5625 $\mu m$ per pixel, while the CSAFE scans are available at a resolution of  0.645 $\mu m$ per pixel.
The length of an average bullet land from Hamby (9 mm Ruger P85) is about 2 millimeter, resulting in signatures of about 1200 pixels for NIST scans, and about 3000 pixels for CSAFE scans.

In comparison, scans from the profilometer used by \citet{chumbley, hadler} were taken at a resolution of about 0.73 $\mu m$ per pixel. The screw driver toolmarks are  about 7 mm in length \citep{manytoolmarks1}, for a total of over 9000 pixels for the width of these scans.

This severe limitation in the amount of available data poses the main challenge in adapting the Chumbley score to matching bullet lands, because of the resulting loss in power.

@@

<!--Markings made by screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length. The Ruger pistol barrels used in the  Hamby study have a 9mm caliber. This means that bullets used in these pistols have land markings of only about a quarter of the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising the problem of potential loss in power of the test. -->

\hh{The Chumbley score allows a separation of a toolmark into raw and normalized digitized versions. Originally, this mechanism is intended to separate between class characteristics and individual markings, however, in the setting of matching bullet striae, we could also use it to separate bullet curvature in profiles from signatures before matching signatures. }

### 2018-03-16 doc/paper.tex
up to chumbley score

Comparisons of striae from bullets are usually based on comparisons of
striae in land engraved areas, which are extracted in form of cross
sections, called \emph{profiles} \citep{aoas,ma2004}. From profiles
bullet \emph{signatures} \citep{chu2013,aoas} are extracted as residuals
of a loess fit (or Gaussian filter). They are considered to be noise
free and a good reflection of the key attributes of the raw marking, and
the unique features of a bullet.
\hh{Do you have a reference for the previous sentenced?} A detailed
discussion of the extraction technique for signatures is given in
\citet{aoas}.

\hh{don't split the discussion on the size. between the next paragraph and }
The scans for the Hamby study come from two sources namely, NIST scans
\citep{nist} and scans produced at Iowa State University
\hh{Iowa State is too general. it's the High-resolution microscopy lab and we are CSAFE}.
Both scans were generated using confocal light microscopes at a 20x
magnification but the microscopes collected data at different
resolutions. The NIST database contains scans of the bullet sets Hamby
44 and 252 at a resolution of 1.5625 \(\mu m\) per pixel. The Iowa State
scans are available for only the Hamby 44 bullet set and at a resoultion
of 0.645 \(\mu m\) per pixel. The data used by \citet{chumbley, hadler}
on the other hand, was produced by a profilometer at a resolution of
about 0.73 \(\mu m\) per pixel. The different resolutions indicate that
the length of the digitized markings would be significantly different
for each setting.

@@

Markings made by screw driver tips \citep{chumbley} are longer and
pronounced and about 7 mm in length \citep{manytoolmarks1}, while the
groove to groove length of the bullet lands from the Hamby study
\citep{hamby} are about 2mm in length. The Ruger pistol barrels used in
the Hamby study have a 9mm caliber. This means that bullets used in
these pistols have land markings of only about a quarter of the size of
screwdriver toolmarks used by \citet{chumbley, hadler}, therefore
raising a problem of distinctiveness. One way to investigate this is to
evaluate bullet striations at different resolutions. This gives us an
incentive to use both the NIST and Iowa State scans and investigate the
effect of length of the digitized mark on the working of the algorithm.
Apart from this, the algorithm developed by \citet{hadler} has only been
tested for flatter and wider surfaces which have negligible curvature,
whereas striations on bullets are made on their curved surfaces.
Therefore, using methods proposed for toolmarks need adaptation in order
to give tangible results for bullets. Another challenge that arises from
these topographical differences, is the understanding of that effect of
different levels of smoothing. Especially investigating its effect on
the algorithm performance when signatures are being compared and when
raw markings or profiles are being compared.

Some other challenges associated with the adaptation process call for a
brief understanding of the methodology involved. The Chumbley method and
algorithm works on the premise of comparison windows that are supposed
to represent small segments of pixels of equal sizes in the two
markings. These windows are to be compared in a predetermined fashion
and are termed as windows of optimization and validation. The
optimization window is selected in the first step called the
optimization step where areas/windows of best agreement between the the
two markings are found.

The window of optimization for bullet striations are shorter as bullet
signatures in terms of absolute length (in mm), are smaller compared to
toolmarks. Even in terms of digitized pixels, there is no
straightforward way of establishing what window size is to be chosen.
One important way to address this is to keep the number of windows of
optimization sufficiently large or not too small. This means, if the
number of windows are too small we inadvertently use bigger window sizes
as we want to compare all segements of one mark with the segments of the
other mark. This leads to not giving enough emphasis to smaller
segements of the two marking that might be in high agreement. In terms
of absolute length we end up comparing smaller segments between the two
markings while dealing with bullet lands, than those being compared for
toolmarks.

For too large window sizes, the weight of small features that would
otherwise uniquely classify a signature and hence identify the region of
agreement is vastly reduced. On the other hand, this introduces another
problem as, if we go too small in the window of optimization, the unique
features of the trace segments are lost and all seem similar. The
problem of distinctiveness in the structure of bullet striations may
also lead to potential failure of the algorithm to identify correct
windows of agreements or not be able to identify these windows at all.
This may lead to incorrect classification of bullets as coming from the
same-source (or matches) or from different source (or non-matches).

Thus, the comaparison windows have a direct influence on the false
positives and false negatives, making identification of the optimum
sizes of these windows very important. This raises important questions
about how and what parameter settings need to be chosen for bullet land
comparison. Something similar is done in \citet{afte-chumbley} for
toolmark comparisons of slip-joint pliers where optimum window sizes are
determined. There is a need to figure out the best parameter settings
which minimizes the errors for unsmoothed markings or profiles and
pre-processed signatures. An analysis of these error rates and
comparison with other methods will help us understand the adaptibility
of the chumbley score to bullets.

>>>>

Comparisons of striae from bullets are usually based on comparisons of
striae in land engraved areas, which are extracted in form of cross
sections, called \emph{profiles} \citep{aoas,ma2004}. From profiles
bullet \emph{signatures} \citep{chu2013,aoas} are extracted as residuals
of a loess fit or Gaussian filter. Signatures are considered to be noise
free and a good reflection of the key attributes of the raw marking, and
the unique features of a bullet.
\hh{Do you have a reference for the previous sentenced?} A detailed
discussion of the extraction technique for signatures is given in
\citet{aoas}.

\hh{don't split the discussion on the size. between the next paragraph and }
There are two sources of scans for sets from the Hamby study available
to us: scans of Hamby 44 and Hamby 252 are available from the NIST
database \citep{nist}. Hamby 44 has also been made available to us and
has been scanned locally for CSAFE at the Roy J.~Carver High Resolution
Microscopy Facility using a Sensofar confocal light microscope. Scans in
the NIST database are made with a NanoFocus at 20x magnification. The
resolutions of the two instruments are different: the NIST scans are
taken at a resolution of 1.5625 \(\mu m\) per pixel, while the CSAFE
scans are available at a resolution of 0.645 \(\mu m\) per pixel. The
length of an average bullet land from Hamby (9 mm Ruger P85) is about 2
millimeter, resulting in signatures of about 1200 pixels for NIST scans,
and about 3000 pixels for CSAFE scans.

In comparison, scans from the profilometer used by
\citet{chumbley, hadler} were taken at a resolution of about 0.73
\(\mu m\) per pixel. The screw driver toolmarks are about 7 mm in length
\citep{manytoolmarks1}, for a total of over 9000 pixels for the width of
these scans.

This severe limitation in the amount of available data poses the main
challenge in adapting the Chumbley score to matching bullet lands,
because of the resulting loss in power.

@@

\hh{The Chumbley score allows a separation of a toolmark into raw and normalized digitized versions. Originally, this mechanism is intended to separate between class characteristics and individual markings, however, in the setting of matching bullet striae, we could also use it to separate bullet curvature in profiles from signatures before matching signatures. }
